---
title: "Risk Preferences Problem Set"
author: "Jeffrey Naecker"
date: "04/21/2019"
output: html_document
---

## Problem 1

### Part (a)

Consider the following two risky options:

+ Option A: 50% chance of \$20, 50% chance of \$50
+ Option B: 25% chance of \$0,  75% chance of \$50

Draw an appropriate Machina triangle to analyze these two options, and carefully show where the two options are in the triangle.

#### solution start

The lower right corner of the triangle should correspond to the \$0 payoff, the lower left to \$20, and the upper left to \$50.  Option A is at $(0, 0.5)$ and option B is at $(0.25, 0.75)$. 

```{r}
## graph of 1a.
library(ggplot2)

gambles <-
  data.frame(
    label = c("A", "B"),
    p1    = c(0, 0.25),
    p3    = c(0.5, 0.75),
    corner1 = c(1,0),
    corner2 = c(0,1)
  )

ggplot(gambles, aes(x = p1, y = p3)) +
  geom_point() + geom_line(aes(x = corner1, y = corner2)) +
  geom_text(aes(label = label), hjust = 0, vjust = 0) + 
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + #stops intercepts from going past 1
  xlim(0,1) + ylim(0,1) #sets scale for x any y axes

ggplot(gambles, aes(x = p1, y = p3)) +
geom_point() +
geom_text(aes(label = label), hjust = 0, vjust = 0)+
xlim(0,1)+ylim(0,1)+geom_segment(aes(x=0,y=0,xend=0,yend=1))+geom_segment(aes(x=0,y=0,xend=1,yend=0))+geom_segment(aes(x=0,y=1,xend=1,yend=0))
```

### Part (b)

Consider an expected utility decision-maker whose value function is given by $u(x) = x^{\frac{1}{2}}$.  Calculate the utility of the two options, and indicate which one the consumer will choose.


```{r}
## code to calculate utilities
## ideally written as general-use function

 utility <- function(p1, p2, p3, x1, x2, x3){
   u= p1*(x1^(1/2)) + p2*(x2^(1/2)) + p3*(x3^(1/2))
   return(u)
 }
 
 x =utility(p1 = 0, p2 = 0.5, p3 = 0.5, x1 = 0, x2 = 20, x3 = 50)
 x
 y = utility(p1 = 0, p2 = 0.25, p3 = 0.75, x1 = 0, x2 = 0, x3 = 50)
y

# We can calculate the utilities as follows:
# 
# ## \begin{align*}
# ## EU_A &= 0.5 \sqrt{20} + 0.5 \sqrt{50} = 5.77 \\
# ## EU_B &= 0.25 \sqrt{0} + 0.75 \sqrt{50} = 5.30
# ## \end{align*}
# 
# Thus the consumer prefers option A.

```

### Part (c)

Draw and give the slope for the indifference curves for the consumer in the previous part.  

#### solution start

We already know that the indifference curves are straight lines with positive slope.  The slope is given by


```{r}
# option 1: prove straight lines, then plot

# option 2: draw countour plot

p<-ggplot(gambles, aes(x = p1, y = p3)) +
geom_point() +
geom_text(aes(label = label), hjust = 0, vjust = 0)+
xlim(0,1)+ylim(0,1)+geom_segment(aes(x=0,y=0,xend=0,yend=1))+geom_segment(aes(x=0,y=0,xend=1,yend=0))+geom_segment(aes(x=0,y=1,xend=1,yend=0))


p+ geom_abline(intercept=0,slope=1.72, color="red")+geom_abline(intercept=.2,slope=1.72, color="red")+geom_abline(intercept=.4,slope=1.72, color="red")

```

### Part (d)

Consider the following two risky options we saw  earlier:

- Option A: 50% chance of \$20, 50% chance of \$50
- Option B: 25% chance of \$0, 75% chance of $50

Consider a prospect theory decision-maker which preference described by $\alpha = 0.5$, $\lambda = 1.5$, $c_r=10$, and $\gamma  = 0.3$.  Calculate their utility of the two options, and indicate which one they will choose.
	
	
#### solution start	
	
First, note that using the probability weighting function we can find that $\pi(0.25) = 0.42$, $\pi(0.5) = 0.5$, and $\pi(0.75) = 0.58$.  We can then calculate the utilities as follows:

```{r}
# similar to part b
#  PT_A &= 0.5 \sqrt{20-10} + 0.5 \sqrt{50-10} = 4.74 \\
# PT_B &= 0.42 \left [-2 \sqrt{|0-10|}\right ] + 0.58 \sqrt{50-10} =1.01 % should be 1.5 not 2?

## Thus this consumer prefers option A as well.

```