---
title: "Hot Hand Assignment"
author: "Jeffrey Naecker"
date: "2/11/2019"
output: html_document
---

## Basic example

From Gelman blog post

flip the coin 4 times

```{r}
flips <- sample(c("T", "H"), 4, replace = T)
```

what faction of flips immediatly following heads were also heads?

```{r}
mean(flips[which(flips == "H") + 1] == "H", na.rm = T)
```

lets repeat this a bunch of times

```{r}
reps <- 1e4
probs <- rep(NA, reps)

for (i in c(1:reps)) {
  flips <- sample(c("T", "H"), 4, replace = T)
  probs[i] <- mean(flips[which(flips == "H") + 1] == "H", na.rm = T)
}

# calculate average probability
mean(probs, na.rm = T) # there will be some NAs because some sets of 4 flips will have no heads
```

The answer you get should be around 0.4, not 0.5 as we might expect.

detecting streaks programatically is a little harder now

**TODO**

need a function to count streaks
given a sequence of H and T, should return a sequence of same length
where each entry counts the number of immediately preceding H's in a row
eg if sequence is HTHHHH, return 0,1,0,1,2,3
also add ability to calculate streaks of tails
HTHHHH becomes 0, 0, 1, 0, 0, 0

```{r}
streaks <- function(flips, count = "H"){
  ##print(flips[3])
  
  
  output = rep(NA, length(flips))
  output[1] = 0
  output[2] = ifelse(flips[1] == count, 1, 0)
  for(i in 3:length(flips)){
      
      if (flips[i-1] == count){
        output[i] = 1 + output[i-1]
      }
      else{
        output[i] = 0
      }
      
          
      #if (flips[i-1] == flips[i-2] & flips[i-1] == count){
      #  output[i] = 1 + output[i-1] 
      #}
      #else{
      #  output[i] = 0
      #}
    
  }
  return(output)
  ##return output
  ##flips_new <- ""
  ##output <- rle(flips)
  ##if flips == "H" &&
  ## if n = n + 1 then count print count(n-1) +1
    
  }
```


## Exercises

### Exercise 1

Veryify the following claim^[https://statmodeling.stat.columbia.edu/2015/07/09/hey-guess-what-there-really-is-a-hot-hand/]:

> With sequences of length 10, the average streakiness statistic (that is, for each person you compute the conditional probability proportion that he gets heads, conditional on him having just got heads on the previous flip, and then you average this across people), is .445. This is pretty far from .5, given that previous estimates of streak-shooting probability have been in the range of 2 percentage points.

10 flips:

```{r}


for (i in c(1:reps)) {
  flips <- sample(c("T", "H"), 100, replace = T)
  probs[i] <- mean(flipsMean[which(flipsMean == "H") + 1] == "H", na.rm = T) ##modify for 2,3
}

mean(probs, na.rm = T)
```



### Exercise 2

Verify the following claim[^https://statmodeling.stat.columbia.edu/2016/02/18/miller-and-sanjurjo-share-5-tips-on-how-to-hit-the-zeitgeist-jackpot/]:

> Imagine Steph Curry shows up to the original hot hand study of Gilovich, Vallone, and Tversky (1985). Steph will shoot 100 shots. Suppose Steph Curry always has the same probability of success, for example, like a 50/50 fair coin. Notice we are assuming Steph Curry doesn’t get hot. Now calculate (1) Curry’s field goal percentage on shots taken when his previous 3 shots (or more) are makes and (2) Curry’s field goal percentage on shots taken when his previous 3 shots (or more) are misses. What do expect? If you are like most people, your intuition tells you that you should expect his field goal percentage to be the same in each case. This intuition is incorrect. In fact, you should expect Curry’s field goal percentage to be 46 percent after 3 (or more) makes, and 54 percent after 3 (or more) misses. Steph Curry is expected to shoot 8 percentage points better after 3 misses in a row than after three makes in row. 

```{r}
#Assume "H" = make
flipsMean <- sample(c("T", "H"), 100, replace = T)
steph <- function(flips, streaks, count = "H"){
      
      output = rep(NA, length(flips))
      #0 means not a streak of at least 3
      output[1] = 0
      output[2] = 0
      output[3] = 0
      #output[2] = ifelse(flips[1] == count, 1, 0)
      
      for(i in 4:length(flips)){
      
      if (flips[i-1] == count){
        output[i] = 1 + output[i-1]
      }
      else{
        output[i] = 0
      }
      }
      
      streakprob = rep(NA, length(output))
      
      
      for (i in 3: length(output)){
        if (output[i] >= 3){
          output[i] = mean(probs, na.rm = T)
        }
        else{
          #0 means not a streak of at least 3
          output[i] = 0
        }
      }
      return(output)
      
      
     # probs[i] <- mean(flips[which(flips == "H") + 1] == "H", na.rm = T)

      
      #for(i in 3:length(output)){
      #  if (output[i]>
      
      #if (flips[i-1] == flips[i-2] == flips[i-3] == count){
      #  return mean(probs, na.rm = T)
        #output[i] = 1 + output[i-1]
      #}else{
      #  output[i] = 0
      #}

  }

```




### Exercise 3

Use a simulation to replicate Figure 1 from the Econometrica paper[^https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2627354].

```{r}
##will have to change lines to adjust for sequence lengths and probabilities
##takes in probability of getting head(p), streak length(k), and number of shots(n)
#expand.grid
#calculate % of shots they make after streak of that length
#function
#expand.grid(c(1:100), c("a","b"))


streaks <- function(flips, count="H"){
  counts=rep(NA, length(flips))
  counts[1]=0
  counts[2]=ifelse(flips[1]==count, 1,0)
  for(i in 3:length(flips)){
    if (flips[i-1]==count){  
     counts[i]=1+counts[i-1]
  }
  else {
    counts[i]=0}

  }
   return(counts)
}


graph<-function(k,n,p){
  reps<-1e4
  probs <- rep(NA, reps)
  if (p==.5){
    if (k==1){
      for (i in c(1:reps)) {
        steph <- sample(c("M", "H"), n, replace = T)
        probs[i] <- mean(steph[which(streaks(steph, count="H")>=1)] == "H")
    }
  
      mean(probs, na.rm=T)
    }
    else if (k==2){
      for (i in c(1:reps)) {
    steph <- sample(c("M", "H"), n, replace = T)
    probs[i] <- mean(steph[which(streaks(steph, count="H")>=2)] == "H")
      }
      mean(probs, na.rm=T)
      }
    else if (k==3){
       for (i in c(1:reps)) {
        steph <- sample(c("M", "H"), n, replace = T)
        probs[i] <- mean(steph[which(streaks(steph, count="H")>=3)] == "H") 
       }
      mean(probs, na.rm=T)
      }
    else {
      for (i in c(1:reps)) {
        steph <- sample(c("M", "H"), n, replace = T)
        probs[i] <- mean(steph[which(streaks(steph, count="H")>=4)] == "H")
      }
     mean(probs, na.rm=T)}
  }
  else if (p==.75){
    if (k==1){
      for (i in c(1:reps)) {
        steph <- sample(c("M", "H","H","H"), n, replace = T)
        probs[i] <- mean(steph[which(streaks(steph, count="H")>=1)] == "H")
    }
  
      mean(probs, na.rm=T)
    }
    else if (k==2){
      for (i in c(1:reps)) {
    steph <- sample(c("M", "H","H","H"), n, replace = T)
    probs[i] <- mean(steph[which(streaks(steph, count="H")>=2)] == "H")
      }
      mean(probs, na.rm=T)
      }
    else if (k==3){
       for (i in c(1:reps)) {
        steph <- sample(c("M", "H","H","H"), n, replace = T)
        probs[i] <- mean(steph[which(streaks(steph, count="H")>=3)] == "H") 
       }
      mean(probs, na.rm=T)
      }
    else {
      for (i in c(1:reps)) {
        steph <- sample(c("M", "H","H","H"), n, replace = T)
        probs[i] <- mean(steph[which(streaks(steph, count="H")>=4)] == "H")
      }
     mean(probs, na.rm=T)}
  }
  else{
    if (k==1){
      for (i in c(1:reps)) {
        steph <- sample(c("M", "M","M","H"), n, replace = T)
        probs[i] <- mean(steph[which(streaks(steph, count="H")>=1)] == "H")
    }
  
      mean(probs, na.rm=T)
    }
    else if (k==2){
      for (i in c(1:reps)) {
    steph <- sample(c("M", "M","M","H"), n, replace = T)
    probs[i] <- mean(steph[which(streaks(steph, count="H")>=2)] == "H")
      }
      mean(probs, na.rm=T)
      }
    else if (k==3){
       for (i in c(1:reps)) {
        steph <- sample(c("M", "M","M","H"), n, replace = T)
        probs[i] <- mean(steph[which(streaks(steph, count="H")>=3)] == "H") 
       }
      mean(probs, na.rm=T)
      }
    else {
      for (i in c(1:reps)) {
        steph <- sample(c("M", "M","M","H"), n, replace = T)
        probs[i] <- mean(steph[which(streaks(steph, count="H")>=4)] == "H")
      }
     mean(probs, na.rm=T)}
  }
}

```

